---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 

```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data: Describe how the observations in the sample are collected, and the implications of this data collection method on the scope of inference (generalizability / causality).

The dataset contains responses 57,601 responses from a survey conducted by the GSS.According to the GSS,most of the interviews were conducted in-person as it is the preferred method with few exceptions of phone interview. The GSS design is a repeated cross-sectional survey of a nationally representative sample of non-institutionalized adults who speak either English or Spanish, and the respondents are answering questions by selecting answers that best matched to their condition

1.The sample size is less than 10% of the entire population

<br>2.The sample observations are independent which means that the sample proportions are approximately normally distributed

<br>3.There are at least 10 Success/Failures in the sample

<br>4.Condition of CLT for proportions have been met and the sample distribution is approximately normal,which enables us to do an inference


* * *

## Part 2: Research question

We define new variable 'Higher Education' as the following:
<br>if the corresponding highest year of education (educ) >12,the respondent accepted higher higher education 
<br>Question: Is race associated with the higher education rate of the US population? 
<br> I am interested in thie question because of I am part of the 'Other race' and I always wonder if different races will expose to different opportunities for getting college education
* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.
```{r}
gss$college<-gss$educ>12
sum(is.na(gss$college)!=TRUE)
df<-gss

#Remove all N/A values in college 
df<-df %>% filter(!is.na(college))

#Bar plot that shows the count of respondents and the proportion of college education receivers

ggplot(df,aes(race,fill=factor(college)))+geom_bar()+geom_text(aes(label=..count..),stat="count",position=position_stack())
# Interpretation: it seems like the college education receiver proportion are lower in Black and the Other race. 

#Distribution of respondent age in each race group: the older the respondent,the less likely that they've received college education 
ggplot(df,aes(race,age,fill=college))+geom_boxplot()
#  The age distributions of each group centered at the mean.The median age of the respondents who received college education (blue color) are around 40 yrs old.
``` 

* * *

## Part 4: Inference

<br>Since we are comparing the proportion of two varibales and one of them has more than 2 levels,We are going to use Chi-Square independance test. We will check if the test-conditions have been met:
<br>Independence:Sample is independent because
<br>1.We have random sample
<br>2.n<10% of the us population
<br>3.no overlaps between each case under each level

<br>Sample Size 
<br>1.COndition met since we have at least 5 expected cases



<br>Variables:Race (white,Black,Others),College(True,False)
<br>Levels:Race contains more than two levels
```{r}
true_prop<-sum(df$college)/length(df$college)
#expected proportion of college education
true_prop

#expected proportion for each group
group_expected<-true_prop*table(df$race)
group_expected
#each of them have >5 success 
```

Hypothesis Testing Set-up
<br>H0(Null Hypothesis):race and college education proportion are independent:there are no relationship between these two variables
<br>HA:race and college are dependent;the higher education rate might be associated with race  

```{r}
#Load the MASS library so we could use the chisq.test function
library(MASS)

#Create a table for the two variables and levels
df_tabl<-table(df$college,df$race)

#Conduct a chi-square independence test 
chisq.test(df_tabl)
```
Conclusion:
<br>1.The p-value is extremely small (<2.2e-16) and significantly lower than alpha=0.05.

<br>In other words, given there are no associate between race and college education rate,the probability of getting the observed value or higher is <0.05; therefore we rejected the null hypothesis with a 95% confidence interval and states that race is associated with the higher education rate. But at this point,we do not have enough evidence to prove that there is a causal reltionship between race and college attendance rate.
<br>Next steps:we need to conduct further research and explore the relationshop between other variables and college education rate. For example,family income,parent education level can also be considered in further analysis.
